Do More Subdivisions in Wards Increase Voter Turnout? An Analysis of the 2022 Toronto Municipal Election

### 2023-06-03 outputs/paper/toronto_elections_paper.qmd
Updating code to graph the relationship between number of subdivisions and voter turnout

### Turnout by Ward
Moreover, using the cleaned voter statistics data set, I added the voter turnout by ward percentage from the bottom of each ward section to a table I created with each ward number. To view a sample of this data set, please see @tbl-summarized_voter_turnout.

@@

#Voter turnout by ward
#Code referenced from: https://github.com/christina-wei/INF3104-1-Covid-Clinics/blob/main/outputs/paper/covid_clinics.pdf
finalized_voter_statistics |>
  ggplot(aes(x = turnout, y = n, label = ward)) + 
  geom_point() + 
  geom_beeswarm() +
  geom_smooth(method=lm, se=FALSE) + 
  geom_text(hjust=-0.5, vjust=0.5) + 
  labs(
    title = "Number of Subdivisions vs Voter Turnout by Ward",
    x = "Turnout (%)",
    y = "Subdivisions",
    label = "Ward"
  ) +
  theme_minimal()
```

@@

Within the data set used for analysis, there are several potential measurement errors. Firstly, the raw data set listed the total number of subdivisions per ward as "NA", despite other aspects of voter turnout and subdivisions being calculated by Toronto Elections. Secondly, there were gaps in the number and corresponding numbering of subdivisions which were not explained by the City Clerk or Toronto Elections. For example, Ward 1 is listed as having 60 subdivisions, plus Advanced Vote and Mail In Voting locations, but subdivisions 4, 20, 43-46, 50, and 58 are unaccounted for, totally 51 subdivisions. Thirdly, there is a lack of clarity whether the people who voted at Advanced Polls or by Mail In Ballot were removed from the total number of electors by subdivision, since these individuals did not vote at their assigned subdivision. If voters cast their ballot early and were not removed from the total number of electors at their specific subdivision, this could impact how voter turnout by subdivision is calculated.

@@

Moreover, the number of subdivisions and polling location placement alone cannot dictate voter behaviour, as there are many other socio-economic factors which inform why an elector does or does not cast a ballot. The caliber of local candidates and the nature of the election, whether an incumbent is standing for re-election or the seat is open for a new councillor impacts voter turnout should be accounted for in future analysis.

>>>>

### Turnout by Ward
Moreover, using the cleaned voter statistics data set, I added the voter turnout by ward percentage calculated by the Toronto Elections from the bottom of each ward section to a table I created with each ward number. To view a sample of this data set, please see @tbl-summarized_voter_turnout.

@@

#Voter turnout by ward
#Code referenced from: https://github.com/christina-wei/INF3104-1-Covid-Clinics/blob/main/outputs/paper/covid_clinics.pdf
finalized_voter_statistics |>
  ggplot(aes(x = turnout, y = n, label = ward)) + 
  geom_point() + 
  geom_beeswarm() +
  geom_smooth(method=lm, se=FALSE) + 
  geom_text(hjust=-0.5, vjust=0.5) + 
  labs(
    title = "Number of Subdivisions vs Voter Turnout by Ward",
    x = "Turnout (%)",
    y = "Subdivisions",
    label = "Ward"
  ) 
```

@@

Within the data set used for analysis, there are several potential measurement errors. Firstly, the raw data set listed the total number of subdivisions per ward as "NA", despite other aspects of voter turnout and subdivisions being calculated by Toronto Elections. Secondly, there were gaps in the number and corresponding numbering of subdivisions which were not explained by the City Clerk or Toronto Elections. For example, Ward 1 is listed as having 60 subdivisions, plus Advanced Vote and Mail In Voting locations, but subdivisions 4, 20, 43-46, 50, and 58 are unaccounted for, totaling 55 subdivisions. Thirdly, there is a lack of clarity whether the people who voted at Advanced Polls or by Mail In Ballot were removed from the total number of electors by subdivision, since these individuals did not vote at their assigned subdivision. If voters cast their ballot early and were not removed from the total number of electors at their specific subdivision, this could impact how voter turnout by subdivision is calculated.

@@

Moreover, the number of subdivisions and polling location placement alone cannot dictate voter behaviour, as there are many other socio-economic factors which inform why an elector does or does not cast a ballot. The caliber of local candidates and the nature of the election, whether an incumbent is standing for re-election or the seat is open for a new councillor impacts voter turnout and should be accounted for in future analysis.
